Towards Verifying the Bitcoin-S Library

### 2019-06-20 paper/paper.tex
more work on intro

\begin{abstract}
  We try to verify properties of the bitcoin-s library, a Scala
  implementation of parts of the Bitcoin protocol. We use the
  Stainless verifier which supports programs in subset of Scala called
  the \emph{Pure Scala Fragment}. We first try to verify the property
  that regular transactions do not create new money. It turns out that
  there is too much code involved that lies outside of the supported
  fragment to make this feasible. However, in the process we uncover
  and fix a bug in bitcoin-s. We then turn to a much simpler (and less
  interesting) property: that adding zero satoshis to a given amount
  of satoshis yields the given amount of satoshis. Here as well a
  significant part of the relevant code lies outside of the supported
  fragment. However, after a series of equivalent transformations, we
  arrive at code that we successfully verify.

@@

\todo{kai: motivation, bitcoin inflation bug, scala better suited for formal verification, bitcoin-s handling millions of dollars, this is a bit of a case study}

\paragraph{The Stainless Verifier.} Stainless is the successor of the
Leon verifier
\cite{DBLP:conf/ecoop/BlancKKS13,DBLP:conf/pldi/VoirolKK15,DBLP:conf/pldi/BlancK15}
and is developed at EPF Lausanne
\cite{Stainless:documentation,Stainless:github}. It is intended to be
used by programmers without training in formal verification and thus
allows to write specifications in Scala and focusses on counterexample
finding in addition to proving correctness.

The following example from \cite{Stainless:documentation} demonstrates
this. Notice how a precondition is specified using the function
\emph{require} and a postcondition using \emph{ensuring}.

@@

\subsection{The Properties to Verify}

Bitcoin-S is a large project found on GitHub implementing many
functionalities.  In this work we look at two of them, namely the
\nameref{property_1} and the \nameref{property_2} described shortly.
Thus, we go through the relevant code parts to verify those
properties, look at some challenges with Stainless and see why we
should write software with verification in mind.  In order to be able
to verify Bitcoin-S, we have to rewrite parts of it in the subset of
Scala supported by Stainless.

\subsubsection{No-Inflation Property}
\label{property_1}

The decision to verify this property was inspired by the bug found in
Bitcoin Core in September 2018 \cite{cve201817144} which allowed to
spend the same unspent transaction output (UTXO) in a transaction
multiple times.  Hence, new coins could be created out of the air.
Thus, we try to verify the property, that a non-coinbase transaction
cannot generate new coins.  Let's name it the \nameref{property_1}.

@@

\subsubsection{Addition-with-Zero Property}
\label{property_2}

\section{Trying to Verify the No-Inflation Property}
\label{chap:verify_check}

@@

\section{Towards Verifying the Addition-with-Zero Property}
\label{chap:verify_add}

>>>>

\begin{abstract}
  We try to verify properties of the bitcoin-s library, a Scala
  implementation of parts of the Bitcoin protocol. We use the
  Stainless verifier which supports programs in subset of Scala called
  the \emph{Pure Scala Fragment}. We first try to verify the property
  that regular transactions do not create new money. It turns out that
  there is too much code involved that lies outside of the supported
  fragment to make this feasible. However, in the process we uncover
  and fix a bug in bitcoin-s. We then turn to a much simpler (and less
  interesting) property: that adding zero satoshis to a given amount
  of satoshis yields the given amount of satoshis. Here as well a
  significant part of the relevant code lies outside of the supported
  fragment. However, after a series of equivalent transformations we
  arrive at code that we successfully verify.

@@

For software handling cryptocurrency, correctness is clearly crucial.
However, even in very well-tested software such as Bitcoin Core,
serious bugs occur. The most recent example is the bug found in
September 2018 \cite{cve201817144} which essentially allowed to
arbitrarily create new coins. Such software is thus a worthwhile
target for formal verification.

\paragraph{The bitcoin-s Library.} Here, we try to verify two
properties of the bitcoin-s library
\cite{BitcoinS:website,BitcoinS:github}, which implements parts of the
Bitcoin protocol in Scala. In particular, it allows to serialize,
deserialize, sign and validate transactions. The library uses
immutable data structures and algebraic data types but is not written
with formal verification in mind. According to the website, the
library is used in production, ``handling millions of dollars each
day'' \cite{BitcoinS:website}.

\paragraph{The Stainless Verifier.} Stainless is the successor of the
Leon verifier
\cite{DBLP:conf/ecoop/BlancKKS13,DBLP:conf/pldi/VoirolKK15,DBLP:conf/pldi/BlancK15}
and is developed at EPF Lausanne
\cite{Stainless:github}. It is intended to be
used by programmers without training in formal verification and thus
allows to write specifications in Scala and focusses on counterexample
finding in addition to proving correctness.

The following example from the Stainless documentation
\cite{Stainless:documentation} demonstrates this. Notice how a
precondition is specified using the function \emph{require} and a
postcondition using \emph{ensuring}.

@@

\section{Trying to Verify the No-Inflation Property}

Thus, we try to verify the property, that a non-coinbase transaction
cannot generate new coins.  Let's name it the \nameref{property_1}.

@@

@@

\section{Towards Verifying the Addition-with-Zero Property}

\label{chap:verify_add}
